~\\[1in] % hack to put space at top.
\textbf{\Huge Acknowledgments}\\

\noindent 
Firstly, I would like to thank my advisors Prof. Andrew Gelman and Prof. Tian Zheng for their continuous support during my PhD study. They have provided tremendous patience, encouragement, and wisdom in every interaction we have had over the past four years. I am the statistician I am today because of you two.

Additionally, I would like to thank Prof. Sarah Cowan and Prof. David Schiminovich for being the best collaborators and mentors I could have asked for. I learned immensely from our work together, and I feel privileged to have been able to learn more about interesting problems in sociology and astronomy. I would also like to thank Prof. Dave Blei and Prof. John Cunningham for being on my defense committee and widening my statistical perspective with your Graphical Models and Gaussian Processes courses.

I also want to express my gratitude to the entire Columbia Department of Statistics, including the faculty, staff, and my fellow graduate students. In particular, I would like to thank Yang, Tim, Jingjing, Peter, Haolei, and Nick for the great basketball memories; and Denis and Rohit for the countless late night tennis sessions. I would also like to thank Dood and Anthony for assisting me with anything and everything. 

Lastly, I would like to thank my parents for working hard their entire lives and showering me with endless love so that I can pursue and achieve my dreams; Harshil for constantly motivating me to aim higher academically and in life; and Dora for being by my side through every failure and success.